---
title: "Intro to APIs: Interacting with PokeAPI"
author: "Landon Batts"
date: '2023-06-21'
output: 
  github_document:
    toc: true
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a vignette to show how to retrieve data from an API. In order to demonstrate, I will be retrieving data from a Pokemon API displaying data of various topics about Pokemon and it's universe.  

I will be writing functions in order to contact [PokeAPI](https://pokeapi.co/) and return usable data frames that can be used to display summaries.

##Required Packages
In order to create functions to contact and interact with PokeAPI, the following packages are needed:

- [`tidyverse`](https://www.tidyverse.org/packages/) - Family of packages for manipulating and visualizing data
- [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/index.html) - package for dealing with and reading in JSON data in R
- [`httr`](https://cran.r-project.org/web/packages/httr/vignettes/quickstart.html)- package used for contacting the API in order to retrieve data

```{r, message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
```

##API Interaction Functions
In order to actually access and interact with data from the API, I created some functions to make it easier and user-friendly.

### `rawtocharacter`
I created this function to apply the rawToChar function in order to convert the raw JSON data to character variables in order to be in a usable form
```{r}
rawtocharacter <- function(x) {
  rawToChar(x$content)
}
```
### `pokemonapi`
I decided to make a function to call the PokeAPI and be able to extract data into a usable form based on the user's options
```{r, warning=FALSE}
pokemonapi <- function(endpoint){
  pokemon <- GET("https://pokeapi.co/api/v2/pokemon?limit=100000&offset=0") #GET function calls to PokeAPI
  parsed <- fromJSON(rawToChar(pokemon$content)) #used fromJSON function to get parsed data in character form
  pokeurl <- parsed$results #saves urls from parsed results
  pokedata <- lapply(lapply(lapply(pokeurl$url,GET),rawtocharacter),fromJSON) #applies GET function to each url in order to get data from all endpoints
  if(endpoint=="abilities"){
    pokedf <- pokeurl %>% select(url) %>% add_column(pokedata) %>%
  unnest_wider(pokedata,names_repair='minimal') %>% 
  unnest_longer(col=abilities)#returns dataset with url and data for ability for each pokemon
    
    pokedf <- pokedf %>% cbind(ability=as.vector(pokedf$abilities$ability$name)) %>% select(name,ability,base_experience,height,weight)
    #makes variable form dataframe variable in order to select usable columns
  }
  if(endpoint=="moves"){
    pokedf <- pokeurl %>% select(url) %>% add_column(pokedata) %>%
  unnest_wider(pokedata,names_repair='minimal') %>% 
  unnest_longer(col=moves) #returns dataset with url and data for each move for every pokemon from moves endpoint
    
    pokedf <- pokedf %>% cbind(move=as.vector(pokedf$moves$move$name)) %>% select(name,move,base_experience,height,weight)
  }
  if(endpoint=="forms"){
    pokedf <- pokeurl %>% select(url) %>% add_column(pokedata) %>%
  unnest_wider(pokedata,names_repair='minimal') %>% 
  unnest_longer(col=forms) #returns dataset with url and data for each form for every pokemon from form endpoint
    
    pokedf <- pokedf %>% cbind(form=as.vector(pokedf$forms$name)) %>% select(name,form,base_experience,height,weight)
  }
  if(endpoint=="stats"){
    pokedf <- pokeurl %>% select(url) %>% add_column(pokedata) %>%
  unnest_wider(pokedata,names_repair='minimal') %>% 
  unnest_longer(col=stats) #returns dataset with url and data for each stat and their level for every pokemon from stats endpoint
    
    pokedf <- pokedf %>% cbind(basestat=as.vector(pokedf$stats$base_stat),statname=as.vector(pokedf$stats$stat$name)) %>% select(name,statname,basestat,base_experience,height,weight)
  }
  if(endpoint=="types"){
    pokedf <- pokeurl %>% select(url) %>% add_column(pokedata) %>%
  unnest_wider(pokedata,names_repair='minimal')  %>%
  unnest_longer(col=types) #returns dataset with url and data for each type for every pokemon from types endpoint
    
    pokedf <- pokedf %>% cbind(type=as.vector(pokedf$types$type$name)) %>% select(name,type,base_experience,height,weight)
  }
  if(endpoint=="game"){
    pokedf <- pokeurl %>% select(url) %>% add_column(pokedata) %>%
  unnest_wider(pokedata,names_repair='minimal') %>% 
  unnest_longer(col=game_indices) #returns dataset with url and data for each game that the Pokemon are included for every pokemon from the game_indices endpoint
    pokedf <- pokedf %>% cbind(games=as.vector(pokedf$game_indices$version$name)) %>% select(name,games,base_experience,height,weight)
  }
  return(pokedf)
}
```
With this function, the user is able to specify one of 6 potential endpoints available and returns a dataframe with data for each. Also included in the data frame is common statistics for Pokemon including experience, height, and weight.

##Data Exploration
Now that the function to read in a dataframe has been created, we can use this to create summaries and graphs. First, I'll get a dataset for each endpoint in order to have data from each stat in a usable form
```{r,cache=TRUE,warning=FALSE}
abilities <- pokemonapi(endpoint="abilities")
moves <- pokemonapi(endpoint="moves")
forms <- pokemonapi(endpoint="forms")
stat <- pokemonapi(endpoint="stats")
stat <- na.omit(stat)
types <- pokemonapi(endpoint="types")
game <- pokemonapi(endpoint="game")
table(types$type)
#Creates contingency table for amount of pokemon for each type
g <- ggplot(data=types,aes(x=type)) #Creates object for type data that can be used to create plots
g + geom_bar(aes(fill="red")) +
theme(axis.text.x = element_text(angle = 90, hjust = 1),legend.position = "none") + labs(x="Pokemon Type",y="Count",title="Pokemon Type Bar Graph")
#Creates labeled bar graph with red bars of pokemon type
#There is a pretty even spread of pokemon types with water having the most pokemon and ice having the least
types <- types %>% mutate(heightweight=weight/height)
#Creates new variable in types data set for the weight height ratio in order to have a continuous variable for plots
g2 <- ggplot(data=types,aes(x=heightweight))
g2 + geom_histogram(fill="red") + xlim(0,400)+labs(title="Distribution of Weight/Height Ratio", x="Weight/Height Ratio",y="Count")
#From the graph, the distribution appears right skewed as a majority of pokemon are smaller with some larger ones such as snorlax
g2 + geom_boxplot()+ xlim(0,400) + labs(title="Boxplot of Pokemon Weight/Height Ratio",x="Weight/Height Ratio")
#Since there is such a wide range of ratios and the data is skewed, there are many outliers present. It appears that the median ratio is around 30 while the maximum is all the way at nearly 1000
g3 <- ggplot(data=stat,aes(x=basestat,y=base_experience))
g3+geom_point(aes(col=statname)) + labs(title="Base Stat Level vs Base XP", x="Base Stat",y="Base Experience")
#This creates a scatterplot comparing the relationship between the base level of a Pokemons stat with their base level experience
#It appears to be somewhat of a positive relationship as the stonger pokemon tend to have more experience.
#I also grouped by statname to see if there were a difference for each type of stat although all stats seem to increase along with experience as well
g4 <- ggplot(data=stat,aes(x=basestat))
g4 + geom_boxplot() + facet_wrap(~statname) + labs(title="Base Stat level for each Stat Type",x="Base Stat")
#This creates box plots for each type of stat over the level of the base stat for each pokemon.
#Each stat has a similar median with attack being the highest slightly. Although, hp has more outliers and a higher max.
```
